LOSA Fuel Poverty maps

### 2021-09-02 LSOA Fuel poverty.Rmd
Fuel poverty and age relationship

# plotting the change in fuel poverty from 2018-2012 of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = pc_diff)) +
  scale_fill_continuous(name = "Change in Proportions of households fuel poor (%)", low = "green", high = "red") +
  labs(caption = "Winchester")
```
### Over 65 annd Fuel Poverty
```{r Loading in the Census age data}
inFile<-paste0(dataFolder, "census/KS102EW-age-categories.csv")
census_age<-readr::read_csv (inFile)
census_age$'LSOA11CD'<- census_age$'geography code'
head(census_age)
names(census_age)
```
```{r Plotting the fuel poverty in over 65s}
#census_age$"LSOA11CD" <- census_age$"geography code"
Census_age_fp <- merge(lsoa_fuelpov_2019, census_age, by = "LSOA11CD")
names(Census_age_fp)
# select the variables we want for clarity
df <- dplyr::select(Census_age_fp, LSOA11CD, "LA Name", "Proportion of households fuel poor (%)", "Proportion Over 65")

head(df)
# select the places we want here
ggplot(df, aes(x='Proportion Over 65', y=`Proportion of households fuel poor (%)`)) +
  geom_point()

census_age$oldpop <- Census_age_fp$'Over 65s'/Census_age_fp$'Age: All usual residents; measures: Value'
ggplot(Census_age_fp, aes(x=oldpop, y=(`Proportion of households fuel poor (%)`/100)) +
  geom_point()
```

>>>>

# plotting the change in fuel poverty from 2018-2012 of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = pc_diff)) +
  scale_fill_continuous(name = "Change in Proportions of households fuel poor (%)", low = "green", high = "red") +
  labs(caption = "Winchester")
```
```{r Loading in the Census age data}
inFile<-paste0(dataFolder, "census/KS102EW-age-categories.csv")
census_age<-readr::read_csv (inFile)
census_age$'LSOA11CD'<- census_age$'geography code'
census_age$prop_65m <- census_age$`Over 65s`/census_age$`Age: All usual residents; measures: Value`
#head(census_age) lots of variables
names(census_age)
message("Original Proportion Over 65:")
summary(census_age$`Proportion Over 65`)
message("New prop_65m:")
summary(census_age$prop_65m)
```

## 2012 Fuel poverty & Census 2011 data

```{r Plotting the 2012 fuel poverty vs 2011 percent over 65s}
#census_age$"LSOA11CD" <- census_age$"geography code"
fp_LSOA_2012_df$LSOA11CD <- fp_LSOA_2012_df$`LSOA Code`
Census_age_2012fp <- merge(fp_LSOA_2012_df, census_age, by = "LSOA11CD")
# select the variables we want for clarity
df_2012 <- dplyr::select(Census_age_2012fp, LSOA11CD, "LA Name", "Proportion of households fuel poor (%)", "Proportion Over 65", prop_65m)
head(df_2012)
# table(df$`LA Name`)
df_2012$la_name <- df_2012$`LA Name` # make a new variable with an easier name
df_2012$pcFuelPoor <- df_2012$`Proportion of households fuel poor (%)` # same - this is actually a % not a proportion!
# select the places we want here
getSolent <- function(df){
  # assumes we are filtering on la_name
  res <- dplyr::filter(df, la_name == "Basingstoke and Deane"|
                           la_name == "East Hampshire"|
                           la_name == "Eastleigh"|
                           la_name == "Fareham"|
                           la_name == "Gosport"|
                           la_name == "Hart"|
                           la_name == "Havant"|
                           la_name == "New Forest"|
                           la_name == "Portsmouth"|
                           la_name == "Rushmoor"|
                           la_name == "Southampton"|
                           la_name == "Test Valley"|
                           la_name == "Winchester"|
                           la_name == "Isle of Wight"
                           )
  return(res)
}
Solent_2012 <- getSolent(df_2012) # use our function to get just Solent LAs
nrow(Solent_2012)
ggplot(Solent_2012, aes(x= 100 * prop_65m, 
                   y= pcFuelPoor, 
                   colour = la_name)) + # use colour for each LA
  geom_point() +
  geom_smooth() + # adds a smoothed best fit line
  theme(legend.position = "None") + # otherwise we get the colour legend at the side
  labs(x = "% aged > 65, 2011 Census",
       y = "% households in fuel poverty, 2012") +
  facet_wrap(. ~ la_name) # draws a separate plot for each LA (in a different colour)
message("Correlation coeficient across al LSOAs (2012)")
cor(Solent_2012$prop_65m, Solent_2012$pcFuelPoor)
```

## 2019 Fuel poverty & Census 2011 data

```{r Plotting the 2019 fuel poverty vs 2011 percent over 65s}
#census_age$"LSOA11CD" <- census_age$"geography code"
Census_age_2019fp <- merge(lsoa_fuelpov_2019, census_age, by = "LSOA11CD")
names(Census_age_2019fp)
# select the variables we want for clarity
df_2019 <- dplyr::select(Census_age_2019fp, LSOA11CD, "LA Name", "Proportion of households fuel poor (%)", "Proportion Over 65", prop_65m)
head(df_2019)
# table(df$`LA Name`)
df_2019$la_name <- df_2019$`LA Name` # make a new variable with an easier name
df_2019$pcFuelPoor <- df_2019$`Proportion of households fuel poor (%)` # same - this is actually a % not a proportion!
# select the places we want here
Solent_2019 <- getSolent(df_2019) # use our function to get just Solent LAs
nrow(Solent_2019)
ggplot(Solent_2019, aes(x= 100 * prop_65m, 
                   y= pcFuelPoor, 
                   colour = la_name)) + # use colour for each LA
  geom_point() +
  geom_smooth() + # adds a smoothed best fit line
  theme(legend.position = "None") + # otherwise we get the colour legend at the side
  labs(x = "% aged > 65, 2011 Census",
       y = "% households in fuel poverty, 2019") +
  facet_wrap(. ~ la_name) # draws a separate plot for each LA (in a different colour)
message("Correlation coeficient across al LSOAs (2019)")
cor(Solent_2019$prop_65m, Solent_2019$pcFuelPoor)
```
